\documentclass{beamer}
\usepackage[english]{babel}

\author{Albert Heinle, U Waterloo}

\title{Benchmarking with SDEval}

\date{February 28, 2014}

\begin{document}
\begin{frame}
\begin{titlepage}
\end{titlepage}
\end{frame}
\begin{frame}
  \frametitle{What is SDEval?}
  
  \begin{itemize}
    \item Tool for {\textbf{creating}} and {\textbf{running}} benchmarks.
    \item Focus on problems in computer algebra.
    \item Built on top of the Symbolic Data project.
  \end{itemize}
  {\textbf{Goals:}}
  \begin{itemize}
    \item Provide a {\textbf{feasible}}, {\textbf{fair}} and
      {\textbf{automatizable way}} of {\textbf{reproducing}} benchmarks
      published in current research works.
    \item Lead to standards for benchmarking, which are not yet thoroughly
      established and followed in computer algebra compared to other research
      areas (e.\,g.\ SAT-solving, SMT-solving, numerics).
    \item Address particularities of benchmarking in computer algebra (results
      of computations sometimes not unique, big variety of subtopics and
      different problems).
    \item Ease the work for researchers who want to compare their
      implementation to existing ones without the necessity to learn to
      operate those.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Creating a Benchmark with SDEval}
  
  Given
  \begin{itemize}
    \item A computation problem (e.\,g.\ Gr{\"o}bner basis calculation).
    \item A list of computer algebra systems or other programs which are
      supported that provide functionality to solve the computation problem.
    \item A list of entries in the Symbolic Data database that can serve as
      inputs for the given computation problem.
  \end{itemize}
  Then our tools can generate executable code for the given computer algebra
  systems, that will instruct them to compute a solution to the computation
  problem given the inputs from Symbolic-Data.
  
  {\textbf{Furthermore:}}
  \begin{itemize}
    \item Tools are easily extendable using templates for custom computation
      problems and custom programs or other computer algebra systems.
    \item Witten in Python, therefore operating-system independent.
    \item GUI available.
  \end{itemize}
\end{frame}
\begin{frame}
  \frametitle{Running Benchmarks with SDEval}
 
  \begin{itemize}
    \item Based on a specific folder structure, that can be generated using
      SDEval.
    \item Running and monitoring computations:
    \begin{itemize}
      \item Progress view via HTML- or XML-file (latter for automatization).
      \item User can predefine restrictions on time and memory consumption
        (tool automatically cancels processes which violate those
        restrictions).
    \end{itemize}
    \item Designed to be easy to customize (e.\,g.\ exclude or add other
      problem instances or programs to existing benchmark).
    \item Using independent, open source time mechanism for fair evaluation.
    \item Every execution of the benchmark will create separate folder with
      results. Therefore change of evaluations of the benchmark over time
      traceable.
  \end{itemize}
\end{frame}

\end{document}
